Supporting information for ``How environmental drivers of spatial synchrony interact''

### 2023-05-30 Paper/SupMat.Rnw
changes in response to referees, minor revisions

If $\epsilon^{(1)}=(\epsilon^{(1)}_1,\ldots,\epsilon^{(1)}_N)$ and 
$\epsilon^{(2)}=(\epsilon^{(2)}_1,\ldots,\epsilon^{(2)}_N)$
are two $N$-dimensional stationary stochastic processes, we define the \emph{cross-variable spatial synchrony}
(or, simply, the \emph{cross synchrony}) between $\epsilon^{(1)}$ and $\epsilon^{(2)}$ as
\begin{linenomath}
\begin{equation}
\rho_{\epsilon^{(1)} \epsilon^{(2)}} = \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}.
\end{equation}
\end{linenomath}
This is a complex-valued function of frequency or timescale. Why is this interpretable as cross-variable 
spatial synchrony? A quantity representing spatial synchrony should compare values of the processes 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ in one location to values in another location, since spatial synchrony
is about relationships between distinct locations; hence our sum is over $i \neq j$. 
A quantity representing cross-variable synchrony should 
make comparisons between two variables, as opposed to comparisons between the same variable in different locations;
hence we use the cross spectrum between variable $1$ in location $i$ ($\epsilon^{(1)}_i$) and variable 
$2$ in location $j$ ($\epsilon^{(2)}_j$). 
Finally, a quantity representing cross-variable synchrony needs to
include the possibility of time delays in the relationship between the two variables. For instance, 
if $\epsilon^{(2)}_j$ is a time-lagged version of $\epsilon^{(1)}_i$, but the effects of 
$\epsilon^{(2)}_j$ on a biotic process $w_j$ are immediate, whereas the effects of $\epsilon^{(1)}_i$
on $w_i$ are delayed by the same lag, then the effects of $\epsilon^{(1)}_i$ on $w_i$ will be correlated
with the effects of $\epsilon^{(2)}_j$ on $w_j$, i.e., this combination of properties of $\epsilon^{(1)}$
and $\epsilon^{(2)}$ will produce synchrony, in the classic sense of correlation, between the biotic
processes $w_i$ and $w_j$.
Thus we use the cross spectrum, $S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}$, for which the phase angle at 
frequency $f$ measures the time delay between frequency-$f$ oscillations of $\epsilon^{(1)}_i$ and 
frequency-$f$ oscillations of $\epsilon^{(2)}_j$.

@@

\section{Model specification, full details}\label{sect:model_spec}

@@

Now, averaging the off-diagonal entries of both sides of (\ref{Sww_T1}-\ref{Sww_T6}) gives
\begin{linenomath}
\begin{align}
\left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{w_i w_j} &= 
\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(1)}} \\
 &+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(2)}} \\
 &+ \frac{1}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\delta_i \delta_j} \\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} 
  \Rp \left( f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} S_{\epsilon_i^{(1)} \epsilon_j^{(2)}} \right) \label{SumSww_T4}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right)  \sum_{i \neq j} \Rp \left( f_{\mcP^{(1)}}  
 S_{\epsilon_i^{(1)} \delta_j}\right)  \label{SumSww_T5}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} \Rp \left( f_{\mcP^{(2)}}    
 S_{\epsilon_i^{(2)} \delta_j}\right), \label{SumSww_T6}
\end{align}
\end{linenomath}
\noindent where $\Rp(\cdot)$ denotes the real part, and where (\ref{SumSww_T4}-\ref{SumSww_T6})
follows from two observations: that the two summands in the brackets of each of the terms 
(\ref{Sww_T4}-\ref{Sww_T6}) are conjugate-transpose pairs; and that, for any complex matrix, $A$,
$\sum_{i \neq j} (A_{ij} +\conj{A_{ji}}) = \sum_{i \neq j} (A_{ij} +\conj{A_{ij}}) = 2 \sum_{i \neq j} \Rp (A_{ij})$.
Now, making use of the definitions of section \ref{Sect:measuring_sync}, this becomes
\begin{linenomath}
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}
+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}
+ \frac{1}{|f_{\mcB}|^2} \rho_{\delta \delta} \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(1)}} \rho_{\epsilon^{(1)} \delta} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(2)}} \rho_{\epsilon^{(2)} \delta} \right),
\end{align}
\end{linenomath}
which expresses population synchrony in terms of noise synchrony and cross synchrony.
If the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ were measured and the process $\delta$ represents
the aggregate effects of unknown processes, then we can write
\begin{linenomath}
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}} \label{rhoww_T1}\\
&+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}} \label{rhoww_T2}\\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \label{rhoww_Tint} \\
&+ \text{other contributions}.
\end{align}
\end{linenomath}
The term (\ref{rhoww_T1}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(1)}$.
The term (\ref{rhoww_T2}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(2)}$.
The term (\ref{rhoww_Tint}) is contributions to the synchrony of $w$ due to interactions between the Moran effects
of $\epsilon^{(1)}$ and $\epsilon^{(2)}$. The ``other contributions'' may include direct Moran effects of other,
unmeasured environmental factors, as well as interactions between these effects and Moran effects of the 
measured factors $\epsilon^{(1)}$ and $\epsilon^{(2)}$.

\section{Details of data and data cleaning}\label{sect:data}

@@

\section{Details of the theoretical case studies}\label{sect:details_case_studies}

@@

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainDirectMoran_ForMS.pdf}
\caption[All theory constituents for kelp for direct Moran effects terms, kelp lags up to 4 quarters]{Plots of all constituents of 
the terms in theory that pertain to direct Moran effects, for kelp, for the CCA1 region (a-o), the SB region
(p-dd) and the CCA2 region (ee-ss), based on a model that considered 
potential kelp lag effects from up to 4 quarters in the past (see Methods).}\label{fig:directMoran_SM_lag4}
\end{figure}

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainInteractingMoran_ForMS.pdf} 
\caption[All theory constituents for kelp for Moran interaction terms, kelp lags up to 4 quarters]{Plots of 
all constituents of the terms in theory that pertain to interactions between Moran effects, for kelp, for CCA1 region
(a-o), the SB region (p-dd), and the CCA2 region (ee-ss), based on a model that considered 
potential kelp lag effects from up to 4 quarters in the past (see Methods).}\label{fig:indirectMoran_SM_lag4}
\end{figure}

>>>>

We briefly compare our definition of synchrony to other measures and terminology 
which have been used. Perhaps the most commonly 
used measure of synchrony is simple Pearson correlation between time series in different locations. When 
time series from multiple locations are available, pairwise Pearson correlation values have sometimes
been considered individually, often as a function of geographic distance between pairs of sites \citep{BjornstadFalck2001},
or have simply been averaged:
\begin{equation}
\left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \cor(w_i,w_j). \label{eq:classic_sync}
\end{equation}
As explained above, our metric, $\rho_{ww}$, is a timescale-specific generalization of 
\begin{equation}
\left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \cov(w_i,w_j), \label{eq:nontssync}
\end{equation}
which is similar to (\ref{eq:classic_sync}), but weights time series of greater variance
more heavily in the average. The latter definition, (\ref{eq:nontssync}), is better connected
to concepts of regional stability because  
\begin{equation}
\var(\sum_i w_i) = \sum_{i} \var(w_i) + \sum_{i \neq j} \cov(w_i,w_j).
\end{equation}
The quantity $\var(\sum_i w_i)$ and related measures are often used as measures of 
regional temporal instability.

The concept of \emph{phase synchonization} has sometimes been used in studies
of populations fluctuating with a single dominant periodicity [e.g., 
\cite{Blasius1999,Massie2010}]. Phase synchronization refers to the tendency of population
peaks in different places to occur simultaneously, and phase synchronization
can occur independently of whether peak magnitudes are related across locations.
Strong phase synchronization would tend to produce a large value of $\rho_{ww}$ at the
common frequency of oscillation.

The term ``asynchrony'' has been used to refer to either
uncorrelated or independent dynamics in different locations. 
If the processes $w_i$ and $w_j$ are independent, then not only are the classic synchrony quantities
$\cor(w_i,w_j)$ and $\cov(w_i,w_j)$ equal to $0$, but so is the cospectrum $\Rp(w_i w_j)$.
On the other hand, if $\cor(w_i,w_j)=\cov(w_i,w_j)=0$, then the cospectrum may nevertheless be non-zero 
if it is positive for some 
frequencies and negative for others in a manner which causes the opposing quantities to cancel.
Research exploring ``asynchrony'' in a timescale-specific context would therefore have to begin 
by clarifying the term.

The term ``anti-synchrony'' has been used to refer to
two time series which are negatively correlated. 
Standard past uses of the term ``anti-synchrony'' may likewise need to be clarified 
in a timescale-specific context.
A negative correlation, $\cor(w_i,w_j)$, or covariance, $\cov(w_i,w_j)$, 
may result from negative associations on some timescales outweighing positive 
associations on other timescales, a pattern which would be clarified by 
the cospectrum, $\Rp(w_i, w_j)$. Related ideas in a community context
were explored by \cite{Zhao2020}.

We now move on to defining cross-variable synchrony. 
If $\epsilon^{(1)}=(\epsilon^{(1)}_1,\ldots,\epsilon^{(1)}_N)$ and 
$\epsilon^{(2)}=(\epsilon^{(2)}_1,\ldots,\epsilon^{(2)}_N)$
are two $N$-dimensional stationary stochastic processes, we define the \emph{cross-variable spatial synchrony}
(or, simply, the \emph{cross synchrony}) between $\epsilon^{(1)}$ and $\epsilon^{(2)}$ as
\begin{linenomath}
\begin{equation}
\rho_{\epsilon^{(1)} \epsilon^{(2)}} = \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}.
\end{equation}
\end{linenomath}
This is a complex-valued function of frequency or timescale. Why is this interpretable as cross-variable 
spatial synchrony? A quantity representing spatial synchrony should compare values of the processes 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ in one location to values in another location, since spatial synchrony
is about relationships between distinct locations; hence our sum is over $i \neq j$. 
A quantity representing cross-variable synchrony should 
make comparisons between two variables, as opposed to comparisons between the same variable in different locations;
hence we use the cross spectrum between variable $1$ in location $i$ ($\epsilon^{(1)}_i$) and variable 
$2$ in location $j$ ($\epsilon^{(2)}_j$). 
Finally, a quantity representing cross-variable synchrony needs to
include the possibility of time delays in the relationship between the two variables. For instance, 
if $\epsilon^{(2)}_j$ is a time-lagged version of $\epsilon^{(1)}_i$, but the effects of 
$\epsilon^{(2)}_j$ on a biotic process $w_j$ are immediate, whereas the effects of $\epsilon^{(1)}_i$
on $w_i$ are delayed by the same lag, then the effects of $\epsilon^{(1)}_i$ on $w_i$ will be correlated
with the effects of $\epsilon^{(2)}_j$ on $w_j$, i.e., this combination of properties of $\epsilon^{(1)}$
and $\epsilon^{(2)}$ will produce synchrony, in the classic sense of correlation, between the biotic
processes $w_i$ and $w_j$.
Thus we use the cross spectrum, $S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}$, for which the phase angle at 
frequency $f$ measures the time delay between frequency-$f$ oscillations of $\epsilon^{(1)}_i$ and 
frequency-$f$ oscillations of $\epsilon^{(2)}_j$.

@@

\section{Main model specification, full details}\label{sect:model_spec}

@@

Now, averaging the off-diagonal entries of both sides of (\ref{Sww_T1}-\ref{Sww_T6}) gives
\begin{linenomath}
\begin{align}
\left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{w_i w_j} &= 
\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(1)}} \\
 &+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(2)}} \\
 &+ \frac{1}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\delta_i \delta_j} \\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} 
  \Rp \left( f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} S_{\epsilon_i^{(1)} \epsilon_j^{(2)}} \right) \label{SumSww_T4}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right)  \sum_{i \neq j} \Rp \left( f_{\mcP^{(1)}}  
 S_{\epsilon_i^{(1)} \delta_j}\right)  \label{SumSww_T5}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} \Rp \left( f_{\mcP^{(2)}}    
 S_{\epsilon_i^{(2)} \delta_j}\right), \label{SumSww_T6}
\end{align}
\end{linenomath}
\noindent where $\Rp(\cdot)$ denotes the real part, and where (\ref{SumSww_T4}-\ref{SumSww_T6})
follows from two observations: that the two summands in the brackets of each of the terms 
(\ref{Sww_T4}-\ref{Sww_T6}) are conjugate-transpose pairs; and that, for any complex matrix, $A$,
$\sum_{i \neq j} (A_{ij} +\conj{A_{ji}}) = \sum_{i \neq j} (A_{ij} +\conj{A_{ij}}) = 2 \sum_{i \neq j} \Rp (A_{ij})$.
Now, making use of the definitions of section \ref{Sect:measuring_sync}, this becomes
\begin{linenomath}
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}
+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}
+ \frac{1}{|f_{\mcB}|^2} \rho_{\delta \delta} \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(1)}} \rho_{\epsilon^{(1)} \delta} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(2)}} \rho_{\epsilon^{(2)} \delta} \right),
\end{align}
\end{linenomath}
which expresses population synchrony in terms of noise synchrony and cross synchrony.
If the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ were measured and the process $\delta$ represents
the aggregate effects of unknown processes, then we can write
\begin{linenomath}
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}} \label{rhoww_T1}\\
&+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}} \label{rhoww_T2}\\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \label{rhoww_Tint} \\
&+ \text{other contributions}.\label{OCeq}
\end{align}
\end{linenomath}
The term (\ref{rhoww_T1}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(1)}$.
The term (\ref{rhoww_T2}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(2)}$.
The term (\ref{rhoww_Tint}) is contributions to the synchrony of $w$ due to interactions between the Moran effects
of $\epsilon^{(1)}$ and $\epsilon^{(2)}$. The ``other contributions'' may include direct Moran effects of other,
unmeasured environmental factors, as well as interactions between these effects and Moran effects of the 
measured factors $\epsilon^{(1)}$ and $\epsilon^{(2)}$.

\section{Additional perspectives on the theory}\label{sect:theorypersp}

Our model (\ref{main_model_stochproc}) can be reconceptualized as 
\begin{equation}
\mcB w = \epsilon^{(T)} + \delta,\label{eq:model_simple}
\end{equation}
if one is willing to consider ``the environment'' to be 
$\epsilon^{(T)} = \mcP^{(1)} \epsilon^{(1)} + \mcP^{(2)} \epsilon^{(2)}$. 
This simplifies the model, producing some advantages, but requires an abstract
definition of the environment which would be harder to measure and less 
directly interpretable. Though this reconceptualization of our model is not
directly useful for the main research goals of our paper, we nevertheless
explore its benefits and drawbacks here, beginning with benefits.

It is straightforward to derive the population spectral characteristics of (\ref{eq:model_simple})
and a relationship between population synchrony and synchrony of the environment. 
Taking the spectral matrix of both sides of (\ref{eq:model_simple}) gives
\begin{equation}
T(\mcB) S_{ww} T(\mcB)^* = S_{\epsilon^{(T)} \epsilon^{(T)}} + S_{\delta \delta} + S_{\epsilon^{(T)} \delta}
+S_{\delta \epsilon^{(T)}}
\end{equation}
and therefore
\begin{equation}
S_{ww} = \frac{1}{|f_{\mcB}|^2}[S_{\epsilon^{(T)} \epsilon^{(T)}} + S_{\delta \delta} + S_{\epsilon^{(T)} \delta}
+S_{\delta \epsilon^{(T)}}].
\end{equation}
Now averaging over all $i,j$ such that $i \neq j$, we have
\begin{align}
\left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_i w_j} &= \frac{1}{|f_{\mcB}|^2} \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{\epsilon^{(T)}_i \epsilon^{(T)}_j} \\
&+ \frac{1}{|f_{\mcB}|^2} \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{\delta_i \delta_j} \\
&+ \frac{2}{|f_{\mcB}|^2} \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \Rp (S_{\epsilon^{(T)}_i \delta_j}),
\end{align}
which gives
\begin{align}
\rho_{ww} &= \frac{1}{|f_{\mcB}|^2} \rho_{\epsilon^{(T)} \epsilon^{(T)}} \label{eq:main_result_simple_1}\\
&+ \text{other contributions}. \label{eq:main_result_simple_2}
\end{align}
This is a simpler version of our main result (equations (\ref{MT-rhoww_T1to3})-(\ref{MT-rhoww_OC})
of the main text and (\ref{rhoww_T1})-(\ref{OCeq}) of SI section \ref{sect:derivation}),
and ``other contributions'' has the same meaning as was explained below (\ref{OCeq}).

We can also use (\ref{eq:model_simple}) to recover the classic Moran theorem, as follows.
If we impose the same assumptions on our model which were used to derive the original
Moran theorem, i.e., $\epsilon^{(T)}$ is iid through time and $\delta$ is $0$, then 
(\ref{eq:model_simple}) gives
\begin{equation}
S_{ww} = \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)} \epsilon^{(T)}},
\end{equation}
and $S_{\epsilon^{(T)} \epsilon^{(T)}}$ does not depend on frequency, $f$.
For a pair of locations, $i$ and $j$, with $i \neq j$, we get
\begin{align}
S_{w_i w_j} &= \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)}_i \epsilon^{(T)}_j}\\
S_{w_i w_i} &= \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)}_i \epsilon^{(T)}_i}\\
S_{w_j w_j} &= \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)}_j \epsilon^{(T)}_j},
\end{align}
so
\begin{align}
\cor(w_i, w_j) &= \frac{\cov(w_i,w_j)}{\sqrt{\var(w_i)\var(w_j)}} \\
&= \frac{\int S_{w_i w_j} df}{\sqrt{\int S_{w_i w_i} df \int S_{w_j w_j} df}} \\
&= \frac{\frac{1}{|f_{\mcB}|^2} \int S_{\epsilon^{(T)}_i \epsilon^{(T)}_j} df}{\sqrt{\frac{1}{|f_{\mcB}|^2} \int S_{\epsilon^{(T)}_i \epsilon^{(T)}_i} df \frac{1}{|f_{\mcB}|^2} \int S_{\epsilon^{(T)}_j \epsilon^{(T)}_j} df}} \\
&= \frac{\int S_{\epsilon^{(T)}_i \epsilon^{(T)}_j} df}{\sqrt{\int S_{\epsilon^{(T)}_i \epsilon^{(T)}_i} df 
\int S_{\epsilon^{(T)}_j \epsilon^{(T)}_j} df}} \\
&= \frac{\cov(\epsilon^{(T)}_i,\epsilon^{(T)}_j)}{\sqrt{\var(\epsilon^{(T)}_i)\var(\epsilon^{(T)}_j)}} \\
&= \cor(\epsilon^{(T)}_i,\epsilon^{(T)}_j),
\end{align}
which proves the original Moran theorem.

Although (\ref{eq:main_result_simple_1})-(\ref{eq:main_result_simple_2}) is simpler than 
our main result in the main text, and relating our results back to the original Moran 
theorem is a benefit, (\ref{eq:main_result_simple_1})-(\ref{eq:main_result_simple_2})
is of limited use for understanding Moran effects because of the abstract notion of 
the environment that it employs. We explain by way of example. Suppose $\epsilon^{(1)}$ 
represents temperature and $\epsilon^{(2)}$ represents precipitation. Then $\mcP^{(1)}$
encodes the nature of the temperature effect on our populations, and $\mcP^{(2)}$
encodes the precipitation effect. Thus $\mcP^{(1)} \epsilon^{(1)} + \mcP^{(2)} \epsilon^{(2)}$
is an abstract environmental quantity, incorporating two distinct traditional environmental
variables and multiple lags. It would not necessarily be straightforward to interpret this
quantity in a more concrete manner than simply ``the aggregate effect of temperature and 
precipitation'' on the populations. Ecologists are more directly interested in the effects 
of directly measurable environmental 
variables such as temperature and precipitation than they are in the effects of highly derived 
variables like $\mcP^{(1)} \epsilon^{(1)} + \mcP^{(2)} \epsilon^{(2)}$. Furthermore,
interactions between Moran drivers cannot be effectively studied if those drivers 
have been combined into a single unified influence.

Although the simpler formulation (\ref{eq:model_simple}) has its strengths and weaknesses,
it is mathematically equivalent to our original model. Because 
\begin{align}
S_{\epsilon^{(T)} \epsilon^{(T)}} &= T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(1)}} T(\mcP^{(1)})^* \\
&+ T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(2)}} T(\mcP^{(2)})^* \\
&+ T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(2)}} T(\mcP^{(2)})^* \\
&+ T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(1)}} T(\mcP^{(1)})^*,
\end{align}
which gives 
\begin{align}
S_{\epsilon^{(T)} \epsilon^{(T)}} &= |f_{\mcP^{(1)}}|^2 S_{\epsilon^{(1)} \epsilon^{(1)}}\\
&= |f_{\mcP^{(2)}}|^2 S_{\epsilon^{(2)} \epsilon^{(2)}}\\
&= f_{\mcP^{(1)}} \overline{f_{\mcP^{(2)}}} S_{\epsilon^{(1)} \epsilon^{(2)}}\\
&= f_{\mcP^{(2)}} \overline{f_{\mcP^{(1)}}} S_{\epsilon^{(2)} \epsilon^{(1)}},
\end{align}
we can average over $i \neq j$ to see that
\begin{align}
\rho_{\epsilon^{(T)} \epsilon^{(T)}} &= |f_{\mcP^{(1)}}|^2 \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(1)}} \\
&+ |f_{\mcP^{(2)}}|^2 \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(2)}}\\
&+ f_{\mcP^{(1)}} \overline{f_{\mcP^{(2)}}} \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(2)}}\\
&+ f_{\mcP^{(2)}} \overline{f_{\mcP^{(1)}}} \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(1)}}.
\end{align}
Therefore, 
\begin{align}
\rho_{\epsilon^{(T)} \epsilon^{(T)}} &= |f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}\\
&+ |f_{\mcP^{(2)}}|^2 \rho_{\epsilon^{(2)} \epsilon^{(2)}}\\
&+ 2\Rp(f_{\mcP^{(1)}} \overline{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}).
\end{align}
Plugging this into (\ref{eq:main_result_simple_1}) recovers our main result.

\section{Details of the setup of the theoretical case studies}\label{sect:case_studies_setup}

For CaseA, we assumed $p_1^{(1)}=\Sexpr{p1_A[2]}$, and $p_i^{(1)}=0$ for $i \neq 1$ and 
$p_0^{(2)}=\Sexpr{p2_A}$ and $p_i^{(2)}=0$ for $i>0$.
For CaseB, we assumed $p_3^{(1)}=\Sexpr{p1_B[4]}$ while $p_i^{(1)}=0$ for $i \neq 3$ and 
$p_0^{(2)} = \Sexpr{p2_B[1]}$ and $p_i^{(2)}=0$ for $i>0$.
For CaseC, we assumed $p_{1}^{(1)} = \Sexpr{p1_C1[2]}$ while $p_i^{(1)}=0$ for $i \neq 1$, and 
$p_{0}^{(2)} = \Sexpr{p2_C1[1]}$ while $p_i^{(2)}=0$ for $i \neq 0$.
For all theoretical case studies, we used $N=\Sexpr{N_A}$, $b_1=\Sexpr{b_A}$,
and $b_i=0$ for $i>1$.
See SI section \ref{sect:examples_cross_sync} and Fig. \ref{fig:rhosForC} for 
information and parameters associated with the noise for each case study.

\section{Details of kelp and related data, and data cleaning}\label{sect:data}

@@

\section{Mathematical details of results from the theoretical case studies}\label{sect:details_case_studies}

@@

\begin{figure}
\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
\caption[Like main text Fig. \ref{MT-fig:PedagogAlt}, but for slightly altered scenarios]{%\doublespacing 
Similar to Fig. \ref{MT-fig:PedagogAlt}, but for slightly altered scenarios. For Fig. \ref{MT-fig:PedagogAlt},
for simplicity, we assumed that larger values of environmental variables had more positive influences
on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
positive population influences (as in Fig. \ref{MT-fig:PedagogAlt}), 
but that larger values of $\epsilon_i^{(2)}$ have more negative population influences (contrasting
with Fig. \ref{MT-fig:PedagogAlt}). In scenario $\beta 1$, 
$l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario $\beta 2$, as in $\alpha 2$,
the environmental effect alignment measure, $l_n+l_{e1}-l_{e2}$, equals $-\sigma/2$ for $\sigma=20$ the period. 
But Moran interactions are synergistic because the 
negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. 
See the online version for color renderings of all figures.}\label{fig:PedagogAlt2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainDirectMoran_ForMS.pdf}
\caption[Extension of main text Fig. \ref{MT-fig:noise} to CCA2, with additional theoretical components pertaining to direct Moran effects]{Plots 
of all constituents of 
the terms in theory that pertain to direct Moran effects, for kelp, for the CCA1 region (a-o), the SB region
(p-dd) and the CCA2 region (ee-ss), based on a model that considered 
potential kelp lag effects from up to 4 quarters in the past (see Methods).}\label{fig:directMoran_SM_lag4}
\end{figure}

@@

\begin{figure}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf} \\
\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainInteractingMoran_ForMS.pdf} 
\caption[Extension of main text Fig. \ref{MT-fig:noise} to CCA2, with additional theoretical components pertaining to Moran interactions]{Plots of 
all constituents of the terms in theory that pertain to interactions between Moran effects, for kelp, for CCA1 region
(a-o), the SB region (p-dd), and the CCA2 region (ee-ss), based on a model that considered 
potential kelp lag effects from up to 4 quarters in the past (see Methods).}\label{fig:indirectMoran_SM_lag4}
\end{figure}
